---
title: "Getting started with wickra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with wickra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`wickra` exposes the Wickra technical-analysis library in R over its C ABI hub.
Every indicator is a constructor returning a `wickra_indicator`; you feed it data
one observation at a time with `update()` — a streaming step whose cost does not
grow with how much history you have already fed it — or run a whole series at
once with `batch()`. Both paths share the exact same Rust core, so a
live feed and a historical backtest compute identical values.

```{r}
library(wickra)
```

## A sample series

The package ships a small synthetic OHLCV series, `sample_ohlcv`, for examples
(a seeded random walk — not real market data).

```{r}
head(sample_ohlcv)
```

## Batch: a whole series at once

Scalar indicators run over a vector with `batch()`. Warmup positions are `NA`.

```{r}
sma <- Sma(20)
sma_values <- batch(sma, sample_ohlcv$close)
tail(sma_values)
```

`batch_fast()` is the opt-in fast form: a SIMD kernel may reassociate the
arithmetic, so it agrees with `batch()` to within a few units in the last place
rather than bit for bit, with `NA` in the same places.

```{r}
fast_values <- batch_fast(Sma(20), sample_ohlcv$close)
max(abs(fast_values - sma_values), na.rm = TRUE) < 1e-9
```

## Streaming: one observation at a time

The same indicator fed tick-by-tick with `update()` returns the identical
values — an equivalence the test suite enforces for every indicator.

```{r}
sma_stream <- Sma(20)
streamed <- vapply(sample_ohlcv$close, function(p) update(sma_stream, p), numeric(1))

same_warmup <- all(is.na(streamed) == is.na(sma_values))
same_values <- all(streamed == sma_values, na.rm = TRUE)
c(batch_equals_streaming = same_warmup && same_values)
```

A typical streaming loop reacts to each value as it arrives:

```{r}
rsi <- Rsi(14)
overbought_days <- 0L
for (price in sample_ohlcv$close) {
  v <- update(rsi, price)            # NA during warmup
  if (!is.na(v) && v > 70) overbought_days <- overbought_days + 1L
}
overbought_days
```

## Multi-output indicators

Indicators with several outputs return a *named* numeric vector (`NA` while
warming up). MACD is the classic example — line, signal, and histogram:

```{r}
macd <- MacdIndicator(12, 26, 9)
last_macd <- c(macd = NA, signal = NA, histogram = NA)
for (price in sample_ohlcv$close) last_macd <- update(macd, price)
last_macd
```

## Candle indicators

Indicators that need the whole bar take the OHLCV fields plus a timestamp:

```{r}
atr <- Atr(14)
last_atr <- NA_real_
for (i in seq_len(nrow(sample_ohlcv))) {
  last_atr <- update(
    atr,
    sample_ohlcv$open[i], sample_ohlcv$high[i], sample_ohlcv$low[i],
    sample_ohlcv$close[i], sample_ohlcv$volume[i], i - 1
  )
}
last_atr
```

## Resetting state

`reset()` returns an indicator to its warmup state so the same object can be
reused on a fresh series:

```{r}
reset(sma)
update(sma, 100)   # NaN — warming up again
```

## Next steps

- Full indicator catalogue, guides, and per-indicator reference:
  <https://docs.wickra.org>.
- Every constructor (`Sma()`, `Rsi()`, `MacdIndicator()`, `Atr()`, …) is listed
  in this package's help index.
